---
title: "Introduction to Quarto"
format: html
bibliography: references.bib
---

# Introduction

## About the data

- Water sample data from Arctic Ocean
- Data downloaded from the Arctic Data Centre [Tweedie ADC Data](https://arcticdata.io/catalog/view/doi:10.18739/A25T3FZ8X)

## Setup

cmd+option+I - adds *inserts chunk* of code (or press the green *+C* button)  
*require* also loads packages like *library*, but quieter

```{r}
#| warning: false

library(readr)
library(here)
```

## Read in data

```{r}
#| warning: false

bg_chem <- read_csv(here("data/BGchem2008data.csv"))
```


# Analysis

## a. Calculate summary statistics

```{r}
#| eval: false
#| echo: false

# Print column names
colnames(bg_chem)

# Show the data structure
str(bg_chem)

# Show the first 6 lines of the data frame
head(bg_chem)

# Get summary of each column (will not calculate for non-numerical values)
summary(bg_chem)

# Print unique values in a specific column
unique(bg_chem$Date)

```


## b. Calculate mean Redfield ratio

```{r}

# Calculate some means
nitrate <- mean(bg_chem$NO3)
nitrite <- mean(bg_chem$NO2)
amm <- mean(bg_chem$NH4)
phos <- mean(bg_chem$P)

```


## Plot Redfield ratio

```{r}
ratio <- (nitrate + nitrite + amm) / phos
```

Here is the ratio we calculated: `r round(ratio)`.


# Conclusion