---
title: "spatial_data"
format: html
---

```{r}
library(readr)
library(sf)
library(ggplot2)
library(leaflet)
library(scales)
library(ggspatial)
library(dplyr)
```

```{r}
knb_url <- "https://dev.nceas.ucsb.edu/knb/d1/mn/v2/object/urn%3Auuid%3Aaceaecb2-1ce0-4d41-a839-d3607d32bb58"

#download the zip file
download.file(url = knb_url, destfile = "../data/shapefile_demo_data.zip")

#unzipping
unzip("../data/shapefile_demo_data.zip", exdir = "../data")

#remove the zip file
file.remove("../data/shapefile_demo_data.zip")

```

## exploring the data using plot() and st_crs()
```{r}
ak_region <- read_sf("../data/ak_regions_simp.shp")

plot(ak_region)
```

```{r}
class(ak_region)

head(ak_region)

glimpse(ak_region)
```

```{r}
st_crs(ak_region)
```

## change CRS with st_transform() - changing to Alaska Albers projection (EPSG:3338)

```{r}
ak_regions_3338 <- ak_region %>%
  st_transform(crs = 3338)

st_crs(ak_regions_3338)
```

```{r}
plot(ak_regions_3338)
```

## sf & dplyr

###select()
```{r}
#return the names of all the columns
colnames(ak_regions_3338)

ak_regions_3338 %>%
  select(region)
```

###filter()
```{r}
unique(ak_regions_3338$region)

ak_regions_3338 %>%
  filter(region == "Southeast")
```

## spatial joins
How many people live in each of these Alaska regions?
```{r}
library(here)
pop_df <- read_csv("../data/alaska_population.csv")

pop_4326_sf <- st_as_sf(pop_df, 
                        coords = c("lng", "lat"),
                        crs = 4326,
                        remove = F)

class(pop_4326_sf)
head(pop_4326_sf)
```

### joining with st_join()

If we were to join the two files as it is, the coordinate systems are different.

```{r}
pop_3338 <- st_transform(pop_4326_sf,
                         crs = 3338)

pop_joined <- st_join(pop_3338,
                      ak_regions_3338,
                      join = st_within)

head(pop_joined)
```
## Calculate total population using group_by() and summarise()

```{r}
pop_region <- pop_joined %>%
  as.data.frame() %>% # will remove geometry, sf -> dataframe
  group_by(region) %>%
  summarise(total_pop = sum(population))

head(pop_region)
class(pop_region)
```

#add total population back tot the shapefile
```{r}
pop_region_3338 <- left_join(ak_regions_3338,
                             pop_region,
                             by = "region")

head(pop_region_3338)

# plot of population by Alaska management area
plot(pop_region_3338["total_pop"])
```

## population by management area
```{r}
pop_mgmt_3338 <- pop_region_3338 %>%
  group_by(mgmt_area) %>%
  summarise(total_pop = sum(total_pop))

plot(pop_mgmt_3338["total_pop"])
```

```{r}
pop_mgmt_3338 <- pop_region_3338 %>%
  group_by(mgmt_area) %>%
  summarise(total_pop = sum(total_pop), do_union = F)

plot(pop_mgmt_3338["total_pop"])
```

## saving saptial object with write_sf()

```{r}
write_sf(pop_region_3338, "../data/ak_regions_populations.shp")
```

## Swithing to using ggplot for mapping

```{r}
ggplot(pop_region_3338) +
  geom_sf(aes(fill = total_pop))
```

```{r}
ggplot(pop_region_3338) +
  geom_sf(aes(fill = total_pop)) +
  labs(fill = "Total Population") +
  scale_fill_continuous(low = "khaki",
                        high = "firebrick",
                        labels = comma)
```

## adding in rivers
```{r}
rivers_3338_sf <- read_sf(here("data/ak_rivers_simp.shp"))

st_crs(rivers_3338_sf)
```

```{r}
ggplot() +
  geom_sf(data = pop_region_3338,
          aes(fill = total_pop))
```

```{r}
ggplot() +
  geom_sf(data = pop_region_3338,
          aes(fill = total_pop)) +
  geom_sf(data = pop_3338,
          size = 0.25) +
  geom_sf(data = rivers_3338_sf,
          aes(linewidth = StrOrder)) + #Strahler stream order (large number = larger river)
  scale_linewidth(range = c(0.025, 0.25), guide = "none") +
  labs(title = "Total Population by Alaska Region",
       fill = "Total Population") +
  scale_fill_continuous(low = "khaki",
                        high = "firebrick",
                        labels = comma) +
  theme_bw()
```

## Adding base maps with ggspatial
 OpenStreetMap uses Pseudo-Mercator (EPSG:3857)
```{r}
pop_3857_sf <- st_transform(pop_3338,
                            crs = 3857)

ggplot(data = pop_3857_sf) +
  ggspatial::annotation_map_tile(type = "osm",
                                 zoom = 4,
                                 progress = "none") +
  geom_sf(aes(color = population),
          fill = NA) +
  scale_color_continuous(low = "darkkhaki",
                         high = "firebrick",
                         labels = comma)
```

# sf objects with leaflet

```{r}
# first need to save projection

epsg3338 <- leaflet::leafletCRS(
  crsClass = "L.Proj.CRS",
  code = "EPSG:3338",
  proj4def = "+proj=aea +lat_0=50 +lon_0=-154 +lat_1=55 +lat_2=65 +x_0=0 +y_0=0 +datum=NAD83 +units=m +no_defs +type=crs",
  resolutions = 2 ^ (16:7)
)
```

## st_transform() to get unprojected coordinate system (WGS 84)
```{r}
pop_region_4326 <- pop_region_3338 %>%
  st_transform(crs = 4326)
```

# make leaflet map

```{r}
m <- leaflet(options = leafletOptions(crs = epsg3338)) %>%
  addPolygons(data = pop_region_4326,
              fillColor = "gray",
              weight = 1)

m
```

# make better map

```{r}
pal <- colorNumeric(palette = "Reds",
                    domain = pop_region_4326$total_pop)
m <- leaflet(options = leafletOptions(crs = epsg3338)) %>%
  addPolygons(
    data = pop_region_4326,
    fillColor = ~pal(total_pop),
    weight = 1,
    color = "black",
    fillOpacity = 0.5,
    label = ~ region
  ) %>%
  addCircleMarkers(
    data = pop_4326_sf,
    lat = ~ lat,
    lng = ~ lng,
    radius = ~ log(population / 500),
    fillColor = "gray",
    fillOpacity = 1,
    weight = 0.25,
    color = "black",
    label = ~ paste0(pop_4326_sf$city, ", population: ", comma(pop_4326_sf$population))
  ) %>%
  addLegend(
      position = "bottomleft",
      pal = pal,
      value = range(pop_region_4326$total_pop),
      title = "Total Population"
    )

m
```

